#First course project for the Reproducible Research course using Knitr
by Francesco Mancino

##Loading in the data
```{r, echo= TRUE}
df<-read.csv("/Users/francescomancino/Desktop/R-programming/Course5/Project1/activity.csv", 
             header = TRUE, sep = ",", na.strings = "NA")
```
##1. What is mean total number of steps taken per day?
```{r, echo=TRUE}
a1<-aggregate(steps ~ date, data= df, sum, na.rm = TRUE)
sm<- mean(a1$steps)
smed<- median(a1$steps)
```
The mean total number of steps taken per day is `r sm` and the median number of steps taken per day is `r smed`.

```{r echo=TRUE}
hist(a1$steps, 
      breaks=seq(from=0, to=35000, by=2500),
      col="blue", 
      xlab="Total number of steps", 
      ylim=c(0, 20), 
      main="Histogram of the total number of steps taken each day with NA's")
```
##2. What is the average daily activity pattern?

The average daily activity is presented in the next plot
```{r echo= TRUE}

a2<-aggregate(steps ~ interval, data= df, mean, na.rm = TRUE)
plot(a2$interval, a2$steps, type = "l")
```

The 5-minute interval that on average across all the days in the data set, contains the maximum number of steps is: 

```{r echo=TRUE}
max_pos <- which(a2$steps == max(a2$steps))
max_interval<- a2[max_pos, 1]
print(max_interval)
```

##3. Imputing missing values

```{r echo = TRUE}
na_count<-sum(is.na(df$steps))
```
There are `r na_count` missing values in the data set.

The strategy for imputing the missing values is presented in the following code chunk:
```{r echo = TRUE}

tidydf<-df
tidydf$steps[is.na(tidydf$steps)] <- median(df$steps, na.rm=TRUE)
```


```{r echo = TRUE}
a3<-aggregate(steps ~ date, data= tidydf, sum)
newmean<- mean(a3$steps)
newmedian<- median(a3$steps)
```

The new mean and median with the imputed values equal `r newmean` and `r newmedian` respectively!

The histogram has changed now that the missing values where replaced by the median number of steps taken each day: 

```{r echo = TRUE}
hist (a3$steps, 
     breaks=seq(from=0, to=35000, by=2000),
     col="green", 
     xlab="Total number of steps", 
     ylim=c(0, 15), 
     main="Histogram of the total number of steps taken each day (with NA imputed)")
```

##4. Are there differences in activity patterns between weekdays and weekends?

```{r echo = TRUE}
tidydf$day=weekdays(as.Date(df$date))
wday =subset(tidydf, ! day %in% c("Saturday","Sunday"))
wkend =subset(tidydf,  day %in% c("Saturday","Sunday"))
tidydf<- rbind(tidydf, wday, wkend)
tidydf$weekend <- as.factor (ifelse(tidydf$day== "Saturday" |
                                tidydf$day== "Sunday", "weekend","weekday"))
```
The previous code prepared the data set with the the sub setting of weekdays and weekends for answering question #4. 

The time series plot for both weekdays and weekends is:
```{r echo = TRUE}
a4<-aggregate(steps ~ interval + weekend, data= tidydf, mean)
library(lattice)
xyplot(steps ~ interval | factor(weekend), data=a4, aspect=1/3, type="l")
```


